A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-01-23 paper/paper.tex
misc

Tables \ref{fulltab:coind}, \ref{fulltab:core}, and \ref{fulltab:frames} are unchanged, except that we make explicit all arguments, including those that can be inferred from the context. The need for transport along a proof of commutation of $\restrf$ in the definition of $\restrl$ is made explicit as shown in table \ref{fulltab:faces}, where the arrow over $\cohf$ indicates the direction of rewrite. The proof of $\cohf$ is described on \ref{fulltab:coh}, and it requires making explicit several equalities which were definitional in extensional type theory. The commutation of $\restrl$ lives in a type referring to $\cohf$, so we need a transport along the commutation of $\restrf$ in the statement of $\cohl$. The proof of $\cohl$ is the most involved proof in the construction. This is where the higher-dimensional coherence condition is needed. The exact formulation of this coherence condition is as follows.
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohc}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\begin{align*}
  \cohf[m][\omega][\theta][r][p][n][p][][d = {\restrf[m][\epsilon][q+2][n][p][][d = \d]}]\;\bullet     \\
  \ap \restrf[m][\omega][r][n][p][][]\;(\cohf[m][\epsilon][\theta][q+1][p][n][p][][d = \d])\;\bullet   \\
  \cohf[m][\epsilon][\omega][q][r][n][p][][d = {\restrf[m][\theta][p][n][p][][d = \d]}] =              \\
  \ap \restrf[m][\theta][p][n][p][][]\;(\cohf[m][\epsilon][\omega][q+1][r+1][n][p][][d = \d])\;\bullet \\
  \cohf[m][\epsilon][\theta][q][p][n][p][][d = {\restrf[m][\omega][r+1][n][p][][d = \d]}]\;\bullet     \\
  \ap \restrf[m][\epsilon][q][n][p][]\;(\cohf[m][\omega][\theta][r][p][n][p][][d = \d])
\end{align*}
where $\mathsf{ap}$ applies a function on two sides of an equality, and $\bullet$ is transitivity of equality. This property of equality proofs holds in \U, and since our formalization is done in \U, the term is trivially discharged.

@@

For a fixed $n$, $\framep$, $\restrf$, and $\cohf$ have to be mutually defined by induction on $p$. Each of them rely on $\painting$, $\restrc$, and $\cohc$ at lower $n$, so they are well-founded. Finally, each of $\painting$, $\restrc$, and $\cohc$ are by induction from $p$ to $n$.

Remark: If we were not working in \U, but in \HGpd, we would need to prove one more higher-dimensional coherence, and if we were working in \Type, we would need to prove arbitrarily many higher-dimensional coherences. For the unfamiliar reader, $\HGpd$ is the subset of types $A$ such that for all $x$ and $y$ in $A$, $x = y$ is in \U. See \cite{herbelin15,altenkirch16,kraus21} for a discussion on the need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory.

>>>>

Tables \ref{fulltab:coind}, \ref{fulltab:core}, and \ref{fulltab:frames} are unchanged, except that we make explicit all arguments, including those that can be inferred from the context. The need for transport along a proof of commutation of $\restrf$ in the definition of $\restrl$ is made explicit as shown in table~\ref{fulltab:faces}, where the arrow over $\cohf$ indicates the direction of rewrite. The proof of $\cohf$ is described on~\ref{fulltab:coh}, and it requires making explicit several rewritings which were invisible in extensional type theory. The commutation of $\restrl$ lives in a type referring to $\cohf$, so we need a transport along the commutation of $\restrf$ in the statement of $\cohl$. The proof of $\cohl$ is the most involved proof in the construction. This is where the higher-dimensional coherence condition is needed. The exact formulation of this coherence condition is as follows.
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohc}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\begin{align*}
  \cohf[m][\omega][\theta][r][p][n][p][][d = {\restrf[m][\epsilon][q+2][n][p][][d = \d]}]\;\bullet     \\
  \ap \restrf[m][\omega][r][n][p][][]\;(\cohf[m][\epsilon][\theta][q+1][p][n][p][][d = \d])\;\bullet   \\
  \cohf[m][\epsilon][\omega][q][r][n][p][][d = {\restrf[m][\theta][p][n][p][][d = \d]}] =              \\
  \ap \restrf[m][\theta][p][n][p][][]\;(\cohf[m][\epsilon][\omega][q+1][r+1][n][p][][d = \d])\;\bullet \\
  \cohf[m][\epsilon][\theta][q][p][n][p][][d = {\restrf[m][\omega][r+1][n][p][][d = \d]}]\;\bullet     \\
  \ap \restrf[m][\epsilon][q][n][p][]\;(\cohf[m][\omega][\theta][r][p][n][p][][d = \d])
\end{align*}
where $\mathsf{ap}$ applies a function on two sides of an equality, and $\bullet$ is transitivity of equality. This property of equality proofs holds in \U, and since our formalization is done in \U, the term is trivially discharged.

@@

For a fixed $n$, $\framep$, $\restrf$, and $\cohf$ have to be mutually defined by induction on $p$. Each of them rely on $\painting$, $\restrc$, and $\cohc$ at lower $n$, so they are well-founded. Finally, each of $\painting$, $\restrc$, and $\cohc$ are built by induction from $p$ to $n$.

Remark: If we were not working in \U, but in \HGpd, we would need to prove one more higher-dimensional coherence, and if we were working in \Type, we would need to prove arbitrarily many higher-dimensional coherences. Here, $\HGpd$ is the subset of types $A$ such that for all $x$ and $y$ in $A$, $x = y$ is in \U. See \cite{herbelin15,altenkirch16,kraus21} for a discussion on the need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory.
